---
title: "AT2 Credit Default"
author: "Edward Truong / Alex Brooks"
date: "26/09/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objectives
1. Develop and deploy a classification model an a product purchase data set.
2. End to end analysis using R
3. Learn the `caret` package for ML
4. Learn to present the case using Rmarkdown

## Read in the dataset

```{r Read in the dataset, error=FALSE, message=FALSE, warning = FALSE}
library(tidyverse)
library(dplyr)      #used by Caret
library(ggplot2)
library(corrplot)

setwd("~/Documents/brookstruong")

import_train <- read_csv("AT2_credit_train_STUDENT.csv")
```

23,101 observations over 17 variables

```{r Trainset, echo=TRUE}
glimpse(import_train)

import_train$default <- as.factor(import_train$default)
```

Looking at the dataset and how the `read_csv` imported the csv, we can understand that the date is imported as a <chr> string rather than <date>. 

```{r Trainset, echo=TRUE}
summary(import_train)
```


```{r Correlation Plot}
cplot <- import_train %>%
  select(-default)

M <- cor(cplot)
corrplot(M, type="upper")
```

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
